{
	\prefacesection{Acknowledgements}
	
	% Single line spacing
	\setstretch{1.0}
	
	It is often said that it is not \emph{what} you know but \emph{who} you know.
	Throughout the course of my PhD I've been exceptionally lucky to have been
	helped along by a great number of people.
	
	Both my supervisor, Jim Garside, and co-supervisor, Steve Furber, have each
	spent countless hours patiently discussing and describing all manner of
	things with me while giving me great freedom in my project. Jim's office door
	has always been open to my unexpected interruptions be it about work, writing
	or walking.  Likewise, Steve has always managed to find time for both
	technical and frivolous endeavours alike. I'm also hugely grateful to Luis
	Plana who has been a rich source of sage advice, insightful questions
	patiently suffered many a foolish question.
	
	Various parts of the work in this thesis (and numerous side projects) would
	not have been possible if not for the multitude of discussions,
	collaborations and even sheer physical hard work of Steve Temple, Javier
	Navaridas, Simon Davidson and Dave Clark. I'm also indebted to Andrew Mundy
	and Jamie Knight, both of whom have donated so much time and effort towards
	verifying and using software implementations of the ideas in this thesis.
	
	The injection of lunchtime silliness by Andrew and Jamie, along with Amanieu
	d'Antras and Andrew Webb and the other CDT members has always brightened my
	day. So to has the friendly and stimulating environment of the School of
	Computer Science and its many staff and students. Of course, I am also very
	grateful for the funding the school has provided for my research.
	
	I cannot thank my wonderful wife, Ann-Marie, enough for being by my side. She
	has given me so much kindness, love and patience and endured a lifetime's
	quota of conversations about hexagons. Finally, thanks too to rest of my
	family, especially my parents, who are to blame for starting me down this
	path and co-suffering drafts and endless rants about this document.
	
	% Required to ensure single line spacing is used for this whole block
	\par%
}
